### 2009-08-28 pkg/mc2d/inst/doc/docmcEnglish.Rnw
Try to provide a True vignette.
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/riskassessment@58 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

\documentclass[10pt,english]{article}
\VignetteIndexEntry
\usepackage[T1]{fontenc}
\usepackage[latin9]{inputenc}
\usepackage[letterpaper]{geometry}
\geometry{verbose,tmargin=2.5cm,bmargin=2.5cm,lmargin=2.5cm,rmargin=2.5cm}
\setlength{\parskip}{\medskipamount}
\setlength{\parindent}{0pt}
\usepackage{url}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{setspace}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% LyX specific LaTeX commands.
%% Because html converters don't know tabularnewline
\providecommand{\tabularnewline}{\\}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Textclass specific LaTeX commands.
\usepackage{noweb}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% User specified LaTeX commands.

\usepackage{babel}

\begin{document}

\title{The {}``\texttt{mc2d}'' package. }

\author{R. POUILLOT, M.-L. DELIGNETTE-MULLER, D.L. KELLY \& J.-B. DENIS}

\maketitle
\tableofcontents{}

\noindent \begin{center}
\vfill{}

\par\end{center}

This documentation is intended for readers with:
\begin{itemize}
\item A medium level of experience in R. Please refer to the Manual {}``An
Introduction to R'' available with R distribution if needed;
\item Some knowledge about Monte-Carlo simulation (its basic principles
and its utility) and about Quantitative Risk Assessment (QRA). 
\end{itemize}
This documentation will not describe all arguments of the functions.
The definitive reference remains the documentation associated with
the package.

\section{Introduction}

\subsection{What is \texttt{mc2d}?}

{}``\texttt{mc2d}'' means Two-Dimensional Monte-Carlo (\emph{{}``Monte-Carlo
à Deux Dimensions''}). This package :
\begin{itemize}
\item provides additional probability distributions;
\item provides tools to construct One-Dimensional and Two-Dimensional Monte-Carlo
Simulations;
\item provides tools to analyse One-Dimensional and Two-Dimensional Monte-Carlo
Simulations.
\end{itemize}
In a previous version, some tools to fit parametric distributions
to data were included. Because these functions are useful for other
purposes, they have been moved to a separate package called \texttt{fitdistrplus}.
Both the \texttt{mc2d} and the \texttt{fitdistrplus} packages are
available at the URL \url{https://r-forge.r-project.org/projects/riskassessment/}.

\texttt{mc2d} was built for QRA in the Food Safety domain but it can
be used in other QRA domains.

\subsection{What is Two-Dimensional Monte-Carlo Simulation (briefly)?}

The following text and Figure \ref{fig:Mc2d} are adapted from \cite{POUILLOT-2004}
and \cite{POUILLOT-2007} where this method was used. The principal
reference for Two-Dimensional Monte-Carlo simulation remains \cite{CULLEN-FREY-1999}.

According to international recommendations, a QRA should reflect the
{}``variability'' in the risk and calculate the {}``uncertainty''
associated with the risk estimate. The {}``variability'' represents
temporal, geographical and/or individual heterogeneity of the risk
for a given population. The {}``uncertainty'' is understood as stemming
from a lack of perfect knowledge about the QRA model structure and
associated parameters%
\footnote{In the engineering risk community, these concepts are refered as \textquotedbl{}aleatoric
uncertainty\textquotedbl{} for \textquotedbl{}variability\textquotedbl{}
and \textquotedbl{}epistemic uncertainty\textquotedbl{} for \textquotedbl{}uncertainty\textquotedbl{}.%
}.

In order to estimate the natural {}``variability'' of the risk,
a Monte-Carlo simulation approach may be useful: the empirical distribution
of the risk within the population may be estimated from the mathematical
combination of distributions reflecting the variability of parameters
across the population.

A two-dimensional (or second-order) Monte-Carlo simulation was proposed
to estimate the {}``uncertainty'' in the risk estimates stemming
from parameter uncertainty \cite{CULLEN-FREY-1999}. A two-dimensional
Monte-Carlo simulation is a Monte-Carlo simulation where the distributions
reflecting \textquotedbl{}variability\textquotedbl{} and the distributions
representing \textquotedbl{}uncertainty\textquotedbl{} are sampled
separately in the simulation, so that \textquotedbl{}variability\textquotedbl{}
and \textquotedbl{}uncertainty\textquotedbl{} in the output may be
estinated separately. It may be described as following (see Figure
\ref{fig:Mc2d}): 
\begin{enumerate}
\item The parameters of the model should be divided into three categories:
the parameters whose distributions reflect {}``variability only'',
hereinafter denoted as {}``variable parameters'', the parameters
whose distributions reflect {}``uncertainty only'', denoted as {}``uncertain
parameters'' and the parameters whose distributions reflect both
uncertainty and variability. For this latter category, a hierarchical
structure, using {}``hyper-parameters'', should be specified: if
a parameter is both uncertain and variable, one should be able to
specify an empirical or parametric distribution representing variability.
This distribution is conditional upon other parameters for which there
is some associated uncertainty. As an example, one should be able
to specify a relationship such as $X\left|a,b\right.\sim N\left(a,b\right)$,
where the specified normal distribution represents variability in
$x$ conditional upon parameters $a$ and $b$. Hyperdistributions,
such as $a\sim Unif\left(l_{a},u_{a}\right)$and $b\sim Unif\left(l_{b},u_{b}\right)$,
represent the uncertainty in the parameters $a$ and $b$;
\item A set of uncertain parameters are randomly sampled from their respective
distributions; 
\item The QRA is performed using a classical (one-dimensional) Monte-Carlo
simulation of size $N_{v}$, treating the uncertain parameters as
fixed. This QRA takes into account the variability in all variable
parameters, and leads to an empirical density function reflecting
the variability of exposure/risk across the population, conditional
upon the uncertain parameters. Various statistics (\emph{e.g.} the
mean, the standard deviation, percentiles) of the resulting empirical
density function are evaluated and stored; 
\item Steps 2) and 3) are performed a large number ($N_{u}$) of times ,
leading to $N_{u}$ sets of statistics; 
\item As output, the $50^{\text{th}}$ percentile (median) of each statistic
is used as a point estimate of this statistic; the $2.5^{\text{th}}$
and $97.5^{\text{th}}$ percentiles of each statistic are used to
establish a 95\% credible interval (CI95) of this statistic. The median
of the $N_{u}$ estimated values for each of the 101 estimated percentiles
allows us to display a {}``variability cumulative distribution''
via a graph. This curve is surrounded by the 2.5th and 97.5th percentiles
obtained from the $N_{u}$ estimates of each of the 101 percentiles.
\end{enumerate}
%
\begin{figure}
\noindent \begin{centering}
\includegraphics[scale=0.5]{mc2dSaumon}
\par\end{centering}

\caption{\label{fig:Mc2d}Shematic Representation of a Two-Dimensional Monte-Carlo
Simulation.}

\end{figure}

\textquotedbl{}\texttt{mc2d}\textquotedbl{} is a set of R functions
that will help to develop such two-dimensional Monte-Carlo simulations.
The main difference from the procedure described above is that \texttt{mc2d}
uses arrays of (at least) two dimensions to derive the results: the
first dimension will reflect variability, the second will reflect
uncertainty. This document will not develop the method further, but
will illustrate the practical application of \texttt{mc2d}, using
a fictitious example.

\subsection{\label{sec:Exemple1}A basic example}

\textbf{Quantitative Risk Assessment:} \textbf{\emph{Escherichia coli}}
\textbf{O157:H7 infection linked to the consumption of frozen ground
beef in <3 year old children.}
\begin{itemize}
\item We assume that, in a given batch of ground beef, \emph{E. coli} O157:H7
are randomly distributed with a mean concentration of $c=10$ bacteria
(cfu) per gram of product;
\item We assume that no bacterial growth occurs in storage, since the product
is kept frozen until it is cooked, just before consumption;
\item 2.7\% of consumers cook their beef {}``rare'', 37.3\% {}``medium''
and 60.0\% {}``well done''; 
\item The following bacterial inactivation $i$ is associated with these
cooking practices:

\begin{itemize}
\item No inactivation for {}``rare'' cooking;
\item 1/5 surviving bacteria for a {}``medium'' cooking;
\item 1/50 surviving bacteria for a {}``well done'' cooking.
\end{itemize}
\item The variability in steak serving sizes $s$ for <3 year children was
estimated in a consumption survey. The {}``best fit'' to the data
was a gamma distribution with parameters: $shape=3.93,\: rate=0.0806$.
\item The dose-response relationship, describing the probability of illness,
$P,$ according to the dose is a one-hit model. The probability of
illness per hit $r$ is assumed to be constant with $r=0.001$.
\end{itemize}
The question is: {}``\emph{What is the distribution of the risk of
illness in the population that consumed the contaminated lot?}''

This distribution will be estimated using Monte-Carlo simulations
performed with R via the {}``\texttt{mc2d}'' package. First, the
model will be developed in a one dimensional framework. Then, in order
to include some uncertainties in the model, it will be derived in
a two dimensional framework.

\subsubsection{One Dimensional Monte-Carlo Simulation}

As a first step, we assume that no uncertainty exists in our model.
All distributions represent variability only. The model may be written
as:\begin{eqnarray*}
c & = & 10.\\
i & \sim & emp\left(\left\{ 1,1/5,1/50\right\} ,\left\{ 0.027,0.373,0.600\right\} \right)\\
s & \sim & gamma\left(3.93,0.0806\right)\\
n & \sim & Poisson\left(c\times i\times s\right)\\
P & = & 1-\left(1-0.001\right)^{n}\end{eqnarray*}

where $emp\left(\boldsymbol{X},\boldsymbol{P}\right)$ is an empirical
distribution wherein each value $X_{i}$ is associated with a probability
$P_{i}$.We will use a \textquotedbl{}classical\textquotedbl{} one
dimensional Monte-Carlo simulation, with 1,000 iterations. Using the
{}``\texttt{mc2d}'' package, the model may be written as:

<<sh0, echo=false>>=
set.seed(666)
@

<<sh1, echo=true>>=
library(mc2d)
ndvar(1000)
conc <- 10
cook <- mcstoc(rempiricalD, values=c(1,1/5,1/50), prob=c(0.027,0.373,0.600))
serving <- mcstoc(rgamma,shape=3.93,rate=0.0806)
expo <- conc * cook * serving
dose <- mcstoc(rpois,lambda=expo)
r <- 0.001
risk <- 1-(1-r)^dose
EC1 <- mc(cook,serving,expo,dose,risk)
print(EC1)
summary(EC1)
@

This One-Dimensional Monte-Carlo simulation provides an estimate of
the mean risk (approximately 5\%), as well as some quantiles of the
risk distribution (2.5\% of the population has a risk of illness greater
than 20.3\%).

\subsubsection{Two dimensional Monte-Carlo Simulation}

Assume now that:
\begin{itemize}
\item The mean concentration of bacteria in the batch is not known with
certainty, but was only a point estimate. Microbiologists think that
the uncertainty around this estimate can be represented via a normal
distribution with parameters $\mu=10$ and $\sigma=2$;
\item Epidemiological studies suggest that the $r$ parameter is also uncertain.
The uncertainty around the mean value of 0.001 can be represented
with a uniform distribution between 0.0005 and 0.0015. 
\end{itemize}
The model could then be written as:\begin{eqnarray*}
c & \sim & N\left(10,2\right)\\
i & \sim & emp\left(\left\{ 1,1/5,1/50\right\} ,\left\{ 0.027,0.373,0.600\right\} \right)\\
s & \sim & gamma\left(3.93,0.0806\right)\\
n & \sim & Poisson\left(c\times i\times s\right)\\
r & \sim & Unif\left(0.0005,0.0015\right)\\
P & = & 1-\left(1-r\right)^{n}\end{eqnarray*}

Note that the distributions of $r$ and $c$ represent uncertainty,
while the distributions of $i$ and $s$ represent variability. $n$,
which is a function of $c$, $i$ and $s,$ will be both variable
\emph{and} uncertain.

We will use a two-dimensional Monte-Carlo simulation, with 1,000 iterations
in the variability dimension and 100 iterations in the uncertainty
dimension. Using the {}``\texttt{mc2d}'' package, the model may
be written as:

<<sh2, echo=true>>=
ndunc(100)
conc <- mcstoc(rnorm,type="U",mean=10,sd=2)
cook <- mcstoc(rempiricalD, type="V",values=c(1,1/5,1/50), prob=c(0.027,0.373,0.600))
serving <- mcstoc(rgamma,type="V",shape=3.93,rate=0.0806)
expo <- conc * cook * serving
dose <- mcstoc(rpois,type="VU",lambda=expo)
r <- mcstoc(runif,type="U",min=0.0005,max=0.0015)
risk <- 1-(1-r)^dose
EC2 <- mc(conc,cook,serving,expo,dose,r,risk)
print(EC2,digits=2)
summary(EC2)
@

Note that the syntax is similar to the earlier model. However, a {}``\texttt{type}''
argument is provided for each distribution, indicating whether the
parameter distribution represents uncertainty (\texttt{type}=\textquotedbl{}\texttt{U}\textquotedbl{}),
variability (\texttt{type}=\textquotedbl{}V\textquotedbl{}), or both
(\texttt{type}=\textquotedbl{}\texttt{VU}\textquotedbl{}).

The summary provides estimates of the variability distributions (in
rows) but with a measure of their uncertainty, linked to the uncertainty
around \texttt{conc} and \texttt{r}. The estimate of the mean risk
is now uncertain. The median of the 100 simulations leads to a \textquotedbl{}best
estimate\textquotedbl{}of 0.0445, with a 95\% \textquotedbl{}credible
interval\textquotedbl{} of {[}0.191, 0.0730{]}.

\section{Basic Principles and Functions}

A typical session of R using {}``\texttt{mc2d}'' is as follows:
\begin{itemize}
\item From data, expert knowledge, \emph{etc.} an empirical or parametric
distribution is chosen for each {}``parent'' parameter. For developing
an empirical distribution from data , the {}``\texttt{fitdistrplus}''
package is recommended;
\item For each parameter, an \texttt{mcnode} object is constructed (key
functions: \texttt{mcdata, mcstoc}); 
\item Various \texttt{mcnode} objects are grouped into an \texttt{mc} object
(key function: \texttt{mc}).
\item The \texttt{mc} object is studied through summaries, graphs, and sensitivity
analysis (key functions: \texttt{summary.mc}, \texttt{plot.mc}, \texttt{tornado},
\texttt{tornadounc}).
\end{itemize}

\subsection{Preliminary Step}

The {}``\texttt{mc2d}'' library should be loaded at the beginning
of your R session (\textquotedbl{}\texttt{library(mc2d)}\textquotedbl{}).

The default size of the Monte-Carlo Simulation should be defined using
the \texttt{ndvar()} function (dimension of variability) and the \texttt{ndunc()}
function (dimension of uncertainty).

\subsection{The \texttt{mcnode} Object as an Elementary Object.}

\subsubsection{\texttt{mcnode} Object Structure}

An \texttt{mcnode} object is the basic element of an \texttt{mc} object.
It is an array of dimension $\left(nsv\times nsu\times nvariates\right)$
where $nsv$ is the dimension of variability, $nsu$ is the dimension
of uncertainty and $nvariates$ is the number of variates of the \texttt{mcnode}%
\footnote{In this section, we will only consider \texttt{mcnode}s with $nvariates=1$.%
}. Four types of \texttt{mcnode} exist: 
\begin{itemize}
\item \textquotedbl{}\texttt{V}\textquotedbl{} \texttt{mcnode}, for \textquotedbl{}\emph{Variability}\textquotedbl{},
is an array of dimension $\left(nsv\times1\times nvariates\right)$.
The distribution represents variability in the parameter; 
\item \textquotedbl{}\texttt{U}\textquotedbl{} \texttt{mcnode}, for \textquotedbl{}\emph{Uncertainty}\textquotedbl{},
is an array of dimension $\left(1\times nsu\times nvariates\right)$.
The distribution represents uncertainty in the parameter. 
\item \textquotedbl{}\texttt{VU}\textquotedbl{} \texttt{mcnode}, for \textquotedbl{}\emph{Variability
and Uncertainty}\textquotedbl{}, is an array of dimension $\left(nsv\times nsu\times nvariates\right)$.
The distribution represents both variability (in the first dimension)
and uncertainty (in the second dimension) in the parameter. 
\item Additionally, a \textquotedbl{}\texttt{0}\textquotedbl{} \texttt{mcnode}
is also defined . \textquotedbl{}\texttt{0}\textquotedbl{} stands
for \textquotedbl{}Neither Variability or Uncertainty\textquotedbl{}.
Such nodes are arrays of dimension $\left(1\times1\times nvariates\right)$.
No uncertainty or variability is considered for these nodes. A \textquotedbl{}\texttt{0}\textquotedbl{}
\texttt{mcnode} is not necessary in the univariate context (use a
scalar instead) but is useful in constructing multivariate nodes (See
section \ref{sec:Multivar}). 
\end{itemize}
%
\begin{figure}
\noindent \begin{centering}
\includegraphics[scale=0.4]{Illustration}
\par\end{centering}

\caption{\label{fig:Structure}Structure of the various \texttt{mcnode} objects.}

\end{figure}

There are 5 ways to construct an \texttt{mcnode} object:
\begin{enumerate}
\item The \texttt{mcstoc} function constructs an \texttt{mcnode} from random
number generating functions;
\item The \texttt{mcdata} function constructs an \texttt{mcnode} from data
sets; 
\item \texttt{An mcnode} can be constructed directly from operations on
\texttt{mcnode} objects;
\item \texttt{mcprobtree} is a special function that constructs an \texttt{mcnode}
from other \texttt{mcnodes} using a probability tree;
\item Some functions, such as {}``\texttt{==}'' or {}``\texttt{>}''
, \texttt{is.na}, \texttt{is.finite} generate a new \texttt{mcnode}
when applied to an existing \texttt{mcnode}.
\end{enumerate}

\subsubsection{\label{sub:mcstoc}The \texttt{\textmd{mcstoc}} function}

The \texttt{mcstoc} function is written as%
\footnote{as is standard in R, most arguments have logical default values and
will be infrequently modified.%
}:

\begin{singlespace}
\noindent \begin{flushleft}
\texttt{mcstoc(func=runif, type=c(\textquotedbl{}V\textquotedbl{},
\textquotedbl{}U\textquotedbl{}, \textquotedbl{}VU\textquotedbl{},
\textquotedbl{}0\textquotedbl{}), ..., nsv=ndvar(), nsu=ndunc(), nvariates=1,
outm=\textquotedbl{}each\textquotedbl{}, nsample=\textquotedbl{}n\textquotedbl{},
seed=NULL, rtrunc=FALSE, linf=-Inf, lsup=Inf, lhs=FALSE)}
\par\end{flushleft}
\end{singlespace}
\begin{itemize}
\item \texttt{func} is a function providing random data or its name as a
character. The table \ref{tab:Distributions} provides available distributions
from the \texttt{stats} and the \texttt{mc2d} libraries that can be
used in \texttt{mcstoc}; 
\item \texttt{type} is the type of requested \texttt{mcnode}. By default,
\texttt{mcstoc} constructs a {}``\texttt{V}'' \texttt{mcnode};
\item \ldots{} are the arguments to be passed to the function \texttt{func},
with the exception of the argument providing the size of the sample.
This latter is calculated by the function according to \texttt{func},
\texttt{type}, \texttt{nsv,} \texttt{nsu} and \texttt{nvariates}.
If the name of the argument specifying the size of the sample is not
\texttt{n} (\emph{e.g.} functions \texttt{rhyper} and \texttt{rwilcox},
see table \ref{tab:Distributions}), the name of this parameter should
be provided in the \texttt{nsample} argument. \emph{Note that all
of the following arguments should be named};
\item \texttt{nsv} and \texttt{nsu} are the number of samples needed in
the variability and uncertainty dimension, respectively. By default,
these values are the ones provided by \texttt{ndvar()} and \texttt{ndunc()},
respectively;
\item \texttt{nvariates} is the desired number of variates in the \texttt{mcnode};
\item \texttt{outm} is the default output for multivariate nodes;
\item \texttt{seed} optionally specifies a seed for the random number generator;
\item \texttt{rtrunc} allows truncation of a distribution between \texttt{linf}
and \texttt{lsup}. This function is not valid for every distribution
(see table \ref{tab:Distributions}). See the \texttt{rtrunc} function
help for further details;
\item \texttt{lhs} allows Latin hypercube sampling of the node . This function
is not valid for every distribution (see table \ref{tab:Distributions}).
See the \texttt{lhs} function help for further details. 
\end{itemize}
%
\begin{table}
\caption{\label{tab:Distributions}Available distributions}

\centering{}\begin{tabular}{|c|c|c|c|c|c|c|}
\hline 
Package & Distribution & function & \multicolumn{1}{c|}{Parameter \texttt{n}} & Other Parameters & trunc & lhs\tabularnewline
\hline
\hline 
\texttt{stats} & beta & \texttt{rbeta} & \texttt{n} & \texttt{shape1, shape2, ncp} & Y & Y\tabularnewline
\hline
 & binomial & \texttt{rbinom} & \texttt{n} & \texttt{size, prob} & Y & Y\tabularnewline
\hline 
 & Cauchy & \texttt{rcauchy} & \texttt{n} & \texttt{location, scale} & Y & Y\tabularnewline
\hline 
 & chi-squared & \texttt{rchisq} & \texttt{n} & \texttt{df, ncp} & Y & Y\tabularnewline
\hline 
 & exponential & \texttt{rexp} & \texttt{n} & \texttt{rate} & Y & Y\tabularnewline
\hline 
 & F & \texttt{rf} & \texttt{n} & \texttt{df1, df2, ncp} & Y & Y\tabularnewline
\hline 
 & gamma & \texttt{rgamma} & \texttt{n} & \texttt{shape, rate (or scale)} & Y & Y\tabularnewline
\hline 
 & geometric & \texttt{rgeom} & \texttt{n} & \texttt{prob} & Y & Y\tabularnewline
\hline 
 & hypergeometric & \texttt{rhyper} & \texttt{nn} & \texttt{m, n, k} & Y & Y\tabularnewline
\hline 
 & lognormal & \texttt{rlnorm} & \texttt{n} & \texttt{meanlog, sdlog} & Y & Y\tabularnewline
\hline 
 & logistic & \texttt{rlogis} & \texttt{n} & \texttt{location, scale} & Y & Y\tabularnewline
\hline 
 & negative binomial & \texttt{rnbinom} & \texttt{n} & \texttt{size, prob (or mu)} & Y & Y\tabularnewline
\hline 
 & normal & \texttt{rnorm} & \texttt{n} & \texttt{mean, sd} & Y & Y\tabularnewline
\hline 
 & Poisson & \texttt{rpois} & \texttt{n} & \texttt{lambda} & Y & Y\tabularnewline
\hline 
 & Student's t & \texttt{rt} & \texttt{n} & \texttt{df, ncp} & Y & Y\tabularnewline
\hline 
 & uniform & \texttt{runif} & \texttt{n} & \texttt{min, max} & Y & Y\tabularnewline
\hline 
 & Weibull & \texttt{rweibull} & \texttt{n} & \texttt{shape, scale} & Y & Y\tabularnewline
\hline
 & Wilcoxon & \texttt{rwilcox} & \texttt{nn} & \texttt{m,n} & Y & Y\tabularnewline
\hline 
\texttt{mc2d} & Bernoulli & \texttt{rbern} & \texttt{n} & \texttt{prob} & Y & Y\tabularnewline
\hline 
 & empirical discrete & \texttt{rempiricalD} & \texttt{n} & \texttt{values, prob} & Y & Y\tabularnewline
\hline 
 & empirical continuous & \texttt{rempiricalC} & n & \texttt{min, max, values, prob} & Y & Y\tabularnewline
\hline
 & PERT & \texttt{rpert} & \texttt{n} & \texttt{min, mode, max, shape} & Y & Y\tabularnewline
\hline
 & triangular & \texttt{rtriang} & \texttt{n} & \texttt{min, mode, max} & Y & Y\tabularnewline
\hline
 & generalised beta & \texttt{rbetagen} & \texttt{n} & \texttt{shape1,shape2,min,max,ncp} & Y & Y\tabularnewline
\hline
 & multinomial & \texttt{rmultinomial} & \texttt{n} & \texttt{n, size, prob} & N & N\tabularnewline
\hline
 & Dirichlet & \texttt{rdirichlet} & \texttt{n} & \texttt{alpha} & N & N\tabularnewline
\hline
 & multivariate normal & \texttt{rmultinormal} & \texttt{n} & \texttt{mean, sigma} & N & N\tabularnewline
\hline
\end{tabular}
\end{table}

In our basic example, \texttt{mcstoc} was used to specify \texttt{conc}
(a normal distribution), \texttt{cook} (an empirical discrete distribution),
\texttt{serving} (a gamma distribution), and \texttt{dose} (a Poisson
distribution). Note that the argument \texttt{lambda} of the Poisson
distribution (node \texttt{dose}) is itself an \texttt{mcnode}.

<<sh3, echo=true, eval=FALSE>>=
conc <- mcstoc(rnorm,type="U",mean=10,sd=2)
cook <- mcstoc(rempiricalD, type="V",values=c(1,1/5,1/50), prob=c(0.027,0.373,0.600))
serving <- mcstoc(rgamma,type="V",shape=3.93,rate=0.0806)
...
dose <- mcstoc(rpois,type="VU",lambda=expo)
r <- mcstoc(runif,type="U",min=0.0005,max=0.0015)
...
@

A normal distribution with parameters $mean=2$, $sd=3$, truncated
on the interval {[}1.5, 2{]}, with samples generated via Latin hypercube
sampling could be written%
\footnote{Note that the mean and the standard deviation of the untruncatednormal
distribution are not preserved in the truncated distribution.%
}:

<<sh3bis, echo=true, eval=true>>=
x <- mcstoc(rnorm, mean=2, sd=3, rtrunc=TRUE, linf=1.5, lsup=2, lhs=TRUE)
summary(x)
@

For convenience in using \texttt{mcstoc}, the following additional
distributions have been implemented: the Bernoulli distribution (\texttt{rbern}),
the empirical discrete distribution (\texttt{rempiricalD}), the PERT
distribution (\texttt{rpert})\cite{VOSE-2000}, the triangular distribution
(\texttt{rtriang}), the Dirichlet distribution (\texttt{rdirichlet})
and the multivariate normal distribution (\texttt{rmultinormal}).
The multinomial distribution has been adapted (vectorized): \texttt{rmultinomial}
(library \texttt{mc2d}) should be used in place of \texttt{rmultinom}
(library \texttt{stats}). The empirical discrete (\emph{e.g.} for
bootstrap), the Dirichlet, the multinomial and the multivariate normal
may be used with uncertain and/or variable parameters by specifying
multivariate nodes. See section \ref{sec:Multivar}.

\subsubsection{\label{sub:mcnode}The \texttt{\textmd{mcdata}} function}

Another way to construct an \texttt{mcnode} object is \emph{via} the
\texttt{mcdata} function, when data are available.

\begin{singlespace}
\noindent \begin{flushleft}
\texttt{mcdata(data, type=c(\textquotedbl{}V\textquotedbl{}, \textquotedbl{}U\textquotedbl{},
\textquotedbl{}VU\textquotedbl{}, \textquotedbl{}0\textquotedbl{}),
nsv=ndvar(), nsu=ndunc(), nvariates=1, outm=\textquotedbl{}each\textquotedbl{})}
\par\end{flushleft}
\end{singlespace}

\noindent \begin{flushleft}
See the documentation associated with this function to see the size/type
of data that can be used to construct an \texttt{mcnode}. The following
example places a \texttt{TRUE} value in a \textquotedbl{}\texttt{U}\textquotedbl{}
node in half of the simulations:
\par\end{flushleft}

<<sh3, echo=true>>=
nu <- ndunc()
tmp <- (1:nu) > (nu/2)
mcdata(tmp,type="U")
@

\subsubsection{Operations on an \texttt{mcnode}}

\texttt{mcnode}s can be automatically constructed using operations
on other \texttt{mcnode}s. Rules are used to transfer uncertainty
and variability coherently within the model. Logically, the rules
are as follows (illustrated here with a {}``+'')%
\footnote{These rules are not the standard R rules for recycling.%
}:
\begin{itemize}
\item \textquotedbl{}\texttt{0}\textquotedbl{} + \textquotedbl{}\texttt{0}\textquotedbl{}
= \textquotedbl{}\texttt{0}\textquotedbl{};
\item \textquotedbl{}\texttt{0}\textquotedbl{} + \textquotedbl{}\texttt{V}\textquotedbl{}
= \textquotedbl{}\texttt{V}\textquotedbl{}
\item \textquotedbl{}\texttt{0}\textquotedbl{} + \textquotedbl{}\texttt{U}\textquotedbl{}
= \textquotedbl{}\texttt{U}\textquotedbl{};
\item \textquotedbl{}\texttt{0}\textquotedbl{} + \textquotedbl{}\texttt{VU}\textquotedbl{}
= \textquotedbl{}\texttt{VU}\textquotedbl{};
\item \textquotedbl{}\texttt{V}\textquotedbl{} + \textquotedbl{}\texttt{V}\textquotedbl{}
= \textquotedbl{}\texttt{V}\textquotedbl{};
\item \textquotedbl{}\texttt{V}\textquotedbl{} + \textquotedbl{}\texttt{U}\textquotedbl{}
= \textquotedbl{}\texttt{VU}\textquotedbl{}: the {}``\texttt{U}''
mcnode is recycled by row, the \textquotedbl{}\texttt{V}\textquotedbl{}
mcnode is recycled in the standard manner by column;
\item \textquotedbl{}\texttt{V}\textquotedbl{} + \textquotedbl{}\texttt{VU}\textquotedbl{}
= \textquotedbl{}\texttt{VU}\textquotedbl{}: the {}``\texttt{V}''
mcnode is recycled in the standard manner by column;
\item \textquotedbl{}\texttt{U}\textquotedbl{} + \textquotedbl{}\texttt{U}\textquotedbl{}
= \textquotedbl{}\texttt{U}\textquotedbl{};
\item \textquotedbl{}\texttt{U}\textquotedbl{} + \textquotedbl{}\texttt{VU}\textquotedbl{}
= \textquotedbl{}\texttt{VU}\textquotedbl{}: the {}``\texttt{U}''
mcnode is recycled by row;
\item \textquotedbl{}\texttt{VU}\textquotedbl{} + \textquotedbl{}\texttt{VU}\textquotedbl{}
= \textquotedbl{}\texttt{VU}\textquotedbl{}
\end{itemize}
Thus, in our example:

<<sh4, echo=true, eval=false>>=
...
expo <- conc * cook * serving
...
risk <- 1-(1-r)^dose
@

\texttt{expo} is a function of a {}``\texttt{U}'' and two {}``\texttt{V}''
\texttt{mcnode}s: it is a {}``\texttt{VU}'' \texttt{mcnode} with
variability in the row dimension and uncertainty in the column dimension
. \texttt{risk} is a function of a {}``\texttt{U}'' and a {}``\texttt{VU}''
\texttt{mcnode}: it is therefore a {}``\texttt{VU}'' \texttt{mcnode}.

\subsubsection{\label{sub:The-mcprobtree-function}The \texttt{\textmd{mcprobtree}}
function}

The \texttt{mcprobtree} function can be used if a {}``probability
tree'' is needed to construct an \texttt{mcnode}. Assume that the
distribution representing the uncertainty on \texttt{conc} was not
itself certain, and that the microbiologists suggest that they are
75\% sure that $conc\sim N\left(10,2\right)$ but that they are 25\%
sure that $conc\sim U\left(8,12\right)$. This could be written using
\texttt{mcprobtree} as%
\footnote{two alternatives for \texttt{whichdist} are \texttt{whichdist <- mcstoc(rempiricalD,
type=\textquotedbl{}U\textquotedbl{}, values=c(0,1), prob=c(75,25))}
or \texttt{whichdist <- mcstoc(rbern,type=\textquotedbl{}U\textquotedbl{},prob=0.25)}%
}:

<<sh4, echo=true, eval=true>>=
conc1 <- mcstoc(rnorm,type="U",mean=10,sd=2)
conc2 <- mcstoc(runif,type="U",min=8,max=12)
whichdist <- c(0.75,0.25)
concbis <- mcprobtree(whichdist,list("0"=conc1,"1"=conc2),type="U")
@

\texttt{mcprobtree} can also be used to generate samples from a mixture
distribution for variability .

\subsubsection{Other functions for constructing an \texttt{mcnode}}

The functions {}``\texttt{==}'', ''\texttt{<}'', ''\texttt{<}='',
''\texttt{>=}'', {}``\texttt{>}'', generate an \texttt{mcnode}
when applied to another \texttt{mcnode}.

Special functions \texttt{is.na(x)}, \texttt{is.nan(x)}, \texttt{is.finite(x)},
\texttt{is.infinite(x)} are implemented to test if any values are
\texttt{NA} (missing data), \texttt{NaN} ({}``\emph{Not A Number}''),
or finite .

<<sh5, echo=true, eval=true>>=
cook < 1
tmp <- log(mcstoc(runif,min=-1,max=1))
tmp
is.na(tmp)
@

\subsubsection{Specifying a correlation between \texttt{mcnodes}}

Structural links between sets of parameters may be very important
in QRA. In \texttt{mc2d}, a Spearman rank correlation structure for
2 or more nodes may be specified with the \texttt{cornode} function.
This function uses the method of Iman \& Conover to generate correlated
samples \cite{IMAN-CONOVER-1982}. Assume that a study suggests that
people who consume rare ground beef alsoconsume larger serving sizes.
We could specify this relation using:

<<sh5bis, echo=true, eval=true>>=
cornode(cook,serving,target=0.5,result=TRUE) 
@

Note that the resulting correlation (around 0.4) is obviously an approximation
to the desired value of 0.5, because a discrete distribution (\texttt{cook}:
3 categories) is correlated with a continuous distribution (\texttt{serving}).

It is possible to create such correlations between {}``V'' nodes,
between {}`` U'' nodes, between {}``VU'' nodes, or between one
{}``V'' node and multiple {}``VU'' nodes.

The use of a multivariate normal distribution (\texttt{rmultinormal})
is another way to specify correlations among nodes, assuming that
the individual nodes are normally distributed.

\subsection{The \texttt{mc} Object}

Once the \texttt{mcnode} objects are constructed, one should group
them into a single object in order to analyse the Monte-Carlo results.
The {}``\texttt{mc}'' object is a list of \texttt{mcnode}s. There
are three ways to construct an \texttt{mc} object: using the \texttt{mc}
function, using the \texttt{evalmcmod} function, or within the \texttt{evalmccut}
function.

\subsubsection{The \texttt{mc} Function}

\noindent \texttt{mc(..., name=NULL, devname=FALSE)}

\ldots{} are \texttt{mcnodes} or \texttt{mc} objects to be gathered
into an \texttt{mc} object. \texttt{mc} value is an \texttt{mc} object
with specific methods, e.g. \texttt{print} or \texttt{summary}. In
our example, we used:

<<sh6, echo=true, eval=false>>=
...
EC2 <- mc(conc,cook,serving,expo,dose,r,risk)
print(EC2)
summary(EC2)
@

\subsubsection{The \texttt{mcmodel} and the \texttt{evalmcmod} Functions}

A model may be written in one step using \texttt{mcmodel} (just a
wrapper of your model in a function), and then evaluated using \texttt{evalmcmod}.
These functions may be used once your model is correct and has been
tested using a small number of iterations. For our example:

<<sh10, echo=true, eval=true>>=
modelEC3 <- mcmodel({
  conc <- mcstoc(rnorm,type="U",mean=10,sd=2)
  cook <- mcstoc(rempiricalD, type="V",values=c(1,1/5,1/50),
    prob=c(0.027,0.373,0.600))
  serving <- mcstoc(rgamma,type="V",shape=3.93,rate=0.0806)
  r <- mcstoc(runif,type="U",min=0.0005,max=0.0015)
  expo <- conc * cook * serving
  dose <- mcstoc(rpois,type="VU",lambda=expo)
  risk <- 1-(1-r)^dose
  mc(conc,cook,serving,expo,dose,r,risk)
})
modelEC3
@

Note that:
\begin{itemize}
\item the model is wrapped between {}``\texttt{\{}'' and {}``\texttt{\}}'';
\item any (valid) R code may be placed in the model%
\footnote{If needed, it is possible to make reference to the simulation dimensions
using \texttt{ndvar()} and/or \texttt{ndunc()}.%
};
\item The model should end with an \texttt{mc()} function\texttt{.}
\end{itemize}
The model is then evaluated using the \texttt{evalmcmod} function:

\noindent \begin{flushleft}
\texttt{evalmcmod(expr, nsv=ndvar(), nsu=ndunc(), seed=NULL)}
\par\end{flushleft}

One can re-run the model with various dimensions or random seeds in
one line:

<<sh11, echo=true, eval=false>>=
EC3 <- evalmcmod(modelEC3,nsv=100,nsu=10,seed=666)
EC4 <- evalmcmod(modelEC3,nsv=100,nsu=1000,seed=666)
@

\subsubsection{The \texttt{mcmodelcut} and the \texttt{evalmccut} Functions}

If evaluating a high-dimensional model, R may exceed its memory limit.
\texttt{evalmccut} evaluates a 2-dimensional Monte-Carlo model (written
with the \texttt{mcmodelcut} function) using a loop, and calculates
and stores statistics in the uncertainty dimension for further analysis.
Readers should refer to the corresponding documentation for further
details. Our example would be written as:

<<sh12, echo=true, eval=FALSE>>=
modEC4 <- mcmodelcut({
{cook <- mcstoc(rempiricalD, type = "V", values = c(0, 1/5, 1/50), 
                prob = c(0.027, 0.373, 0.6))
 serving <- mcstoc(rgamma, type = "V", shape = 3.93, rate = 0.0806)       
 conc <- mcstoc(rnorm, type = "U", mean = 10, sd = 2)       
 r <- mcstoc(runif, type = "U", min = 5e-04, max = 0.0015)     
}
## Second block 
{expo <- conc * cook * serving       
 dose <- mcstoc(rpois, type = "VU", lambda = expo)
 risk <- 1 - (1 - r)^dose
 res <- mc(zero,conc, cook, serving, expo, dose, r, risk)     }
## Third block 
{list(
  sum = summary(res),
  plot = plot(res, draw=FALSE),
  minmax = lapply(res,range),
  tor=tornado(res),
  et =  sapply(res,sd))
}
})
evalmccut(modEC4, nsv = 10001, nsu = 101, seed = 666)
@

Note that the use of a \texttt{tornado} function in the model should
be avoided as it slows the \texttt{evalmccut} function considerably.
The \texttt{tornado} function will be rewritten in the near future
to improve its performance.

\subsection{Analysing an \texttt{mc} Object}

As a reminder, the \texttt{print} function provides a very basic summary
of the \texttt{mc} object. It has a \texttt{digits} argument (default:
3). Obviously, other more informative functions are provided in the
\texttt{mc2d} package.

\subsubsection{The \texttt{summary} Function}

The \texttt{summary} function provides statistics on an \texttt{mc}
object:

\noindent \begin{flushleft}
\texttt{summary(object, probs=c(0,0.025,0.25,0.5,0.75,0.975,1), lim=c(0.025,0.975),
...)}
\par\end{flushleft}

The mean, the standard deviation and the quantiles provided in the
\texttt{probs} arguments are evaluated on the variability dimension.
Then, the median and the quantiles provided in the \texttt{lim} argument
are evaluated on these statistics. Of course, these arguments should
be changed if other quantiles are needed.

<<sh14, echo=true, eval=true>>=
tmp <- summary(EC2,probs=c(0.995,0.999),digits=12)
tmp$risk
@

\subsubsection{The \texttt{hist} Function}

The \texttt{hist} provides a histogram of the different \texttt{mcnodes}
making up the \texttt{mc} object (cf. Figure \ref{fig:Function-hist}).

\noindent \begin{flushleft}
\texttt{hist (x, griddim = NULL, xlab = names(x), ylab = \textquotedbl{}Frequency\textquotedbl{},
main = \textquotedbl{}\textquotedbl{}, ...)}
\par\end{flushleft}

In the current version, uncertainty and variability distributions
are collapsed. Thus, the resulting histogram may be meaningless.

<<sh15, echo=true, eval=false, fig=false>>=
hist(EC2)
@

%
\begin{figure}
\caption{\label{fig:Function-hist}Function \texttt{hist}.}

<<sh16, echo=false, eval=true, fig=true>>=
hist(EC2)
@
\end{figure}

\subsubsection{The \texttt{plot} function}

The \texttt{plot} function provides a graph of the empirical distribution
function of the estimate (mean or median) of the quantiles.

\noindent \begin{flushleft}
\texttt{plot (x, prec = 0.01, stat = c(\textquotedbl{}median\textquotedbl{},
\textquotedbl{}mean\textquotedbl{}), lim = c(0.025,0.975), na.rm =
TRUE, griddim = NULL, xlab = NULL, ylab = \textquotedbl{}Fn(x)\textquotedbl{},
main = \textquotedbl{}\textquotedbl{}, draw = TRUE, ...)}
\par\end{flushleft}

For our example, see Figure \ref{fig:Function-plot}, a default graph.

<<sh17, echo=true, eval=false, fig=false>>=
plot(EC2)
@

%
\begin{figure}
\caption{\label{fig:Function-plot}\texttt{plot} Function .}

<<sh18, echo=false, eval=true, fig=true>>=
plot(EC2)
@
\end{figure}

Note that \texttt{mcnode} objects have the same methods \texttt{print},
\texttt{summary}, \texttt{plot}, and \texttt{hist}.

\subsubsection{The \texttt{tornado} function}

The \texttt{tornado} function calculates the Spearman (default) rank
correlation between nodes of the \texttt{mc} object.

\noindent \begin{flushleft}
\texttt{tornado(x, output=length(x), use=\textquotedbl{}all.obs\textquotedbl{},
method=c(\textquotedbl{}spearman\textquotedbl{}, \textquotedbl{}kendall\textquotedbl{},\textquotedbl{}pearson\textquotedbl{}),
lim=c(0.025, 0.975))}
\par\end{flushleft}

where \texttt{output} is the \texttt{mcnode} (name or rank) of the
output (default: the last \texttt{mcnode}). Missing data are treated
using the \texttt{use} arguments (see the reference documentation).
\texttt{tornado} creates a \texttt{tornado} object with a \texttt{plot}
method (\emph{cf.} Figure \ref{fig:Function-plottor}).

<<sh19, echo=true, eval=true, fig=false>>=
torEC2 <- tornado(EC2)
plot(torEC2)
@

%
\begin{figure}
\caption{\label{fig:Function-plottor}\texttt{plot}.\texttt{tornado} Function
.}

<<sh19b, echo=false, eval=true, fig=true>>=
plot(torEC2)
@
\end{figure}

\subsubsection{The \texttt{tornadounc} function}

The \texttt{tornadounc} function examines the impact of the uncertainty
on the estimate of an output. It calculates the Spearman (default)
rank correlation between statistics of the \texttt{mc} object in the
variability dimension.

\noindent \begin{flushleft}
\texttt{tornadounc(mc,output = length(mc), quant=c(0.5,0.75,0.975),
use = \textquotedbl{}all.obs\textquotedbl{}, method=c(\textquotedbl{}spearman\textquotedbl{},\textquotedbl{}kendall\textquotedbl{},\textquotedbl{}pearson\textquotedbl{}),
...)}
\par\end{flushleft}

The \texttt{quant} argument indicates which quantiles should be used
in the variability dimension. \texttt{tornadounc} creates a \texttt{tornadounc}
object with a \texttt{plot} method

<<sh19c, echo=true, eval=true, fig=false>>=
tornadounc(EC2, output="risk", quant=.99)
@

The output shows the impact of the uncertain nodes (type \textquotedbl{}U\textquotedbl{}
nodes) and some statistics (mean, median and, here, the $99^{\text{th}}$percentile)
calculated on the variability dimension (type \textquotedbl{}V\textquotedbl{}
and type \textquotedbl{}VU\textquotedbl{} nodes) of some output statistics
.

\subsection{Other Functions and \texttt{mc} Objects}

\texttt{mc} objects are simply lists of three dimensional arrays;
within each array, values in a given column represent variability
in the parameter.

Knowing the structure of the \texttt{mc} and the structure of the
\texttt{mcnode} objects, it is straightforward to apply any R function
to these objects. The \textquotedbl{}\texttt{\$}\textquotedbl{} function
is helpful for extracting an \texttt{mcnode} from an \texttt{mc} object.
The \texttt{unmc} function removes all attributes, classes, and dimensions
equal to one, providing a list of vectors, matrices and/or arrays.

Here is a (silly) example building a linear model (in fact \texttt{ndunc()}
linear models) between the \texttt{risk} and the \texttt{dose} within
each uncertainty dimension and estimating some statistics for the
coefficients. This example is here only to illustrate that the entire
spectrum of R functionality is available for your analysis.

<<sh19ter, echo=true, eval=true>>=
tmp <- unmc(EC2, drop=TRUE)
dimu <- ncol(tmp$risk)
coef <- sapply(1:dimu, function(x) lm(tmp$risk[,x] ~ tmp$dose[,x])$coef) 
apply(coef,1,summary)
@

\section{\label{sec:Multivar}Multivariate Nodes}

The dimension \texttt{nvariates} is the third dimension of the \texttt{mcnode}.
One can ignore it while using \texttt{mc2d} . Nevertheless, its use
is mandatory to handle some multivariate distributions, and it may
be useful in other circumstances. Constructing multivariate nodes
is straightforward. We note that the following code:

<<sh19ter, echo=true, eval=false>>=
mcstoc(runif, nvariates=3, min=c(1,2,3),max=4)
@

will logically not provide a node with 3 variates, each having a different
limit. The recycling rule says that \texttt{c(1, 2, 3)} will be used
in the first dimension, i.e. the variability dimension. Use instead:

<<sh19quatr, echo=true, eval=true>>=
lim <- mcdata(c(1,2,3), type="0", nvariates=3)
mcstoc(runif, nvariates=3, min=lim,max=4)
@

\subsection{Multivariate Nodes for Multivariate Distributions}

The basic usage of multivariate nodes (and the reason why they have
been implemented) is for multivariate distributions such as the Dirichlet
distribution, the multinomial distribution, the multivariate normal
distribution and, possibly, the empirical distribution

As an example, assume that 3-member families buy 500 g of ground beef.
The proportions of steak eaten by the baby, his older brother and
his mother follow a Dirichlet (uncertainty) distribution with (vector)
parameter $\alpha=(2,3,5)$. We want to derive the distribution (variability)
of steak eaten by 500 babies sampled from these 500 families.

<<sh20, echo=true, eval=true>>=
(p <- mcstoc(rdirichlet, type="U", nsu=100, nvariates=3, alpha=c(2,3,5)))
s <- mcstoc(rmultinomial,type="VU", nsv=500, nsu=100, nvariates=3,size=500, prob=p)
summary(s)
@

Assume that each member of these families eats a {}``normal'' distribution
(variability) of steak with mean 100, 150 and 250 g. There is a positive
correlation between the servings of the children, and a negative one
with the serving of the mother. We want to derive the distribution
(variability) of steak eaten by 500 babies.

<<sh21, echo=true, eval=true>>=
(x <- mcstoc(rmultinormal,type="V", nvariates=3, mean=c(100,150,250), 
                              sigma=c(10,2,-5,2,10,-5,-5,-5,10))) 
cor(x[,1,])
@

In this example, \texttt{mean} could be variable or uncertain, as
well as \texttt{sigma}%
\footnote{Caution: the use of a varying \texttt{sigma} can make the analysis
very slow.%
}. You could have used, for an uncertain mean:

<<sh21, echo=true, eval=true>>=
m <- mcdata(c(100,150,250), type="0", nvariates=3)
mun <- mcstoc(rnorm, type="U", nvariates=3, mean=m, sd=20)
x <- mcstoc(rmultinormal, type="VU", nvariates=3, mean=mun, sigma=c(10,2,-5,2,10,-5,-5,-5,10))
cor(x[,1,])
@

The correlation is preserved, but the mean of each category is uncertain.

Finally, multivariate nodes may be useful to derive a nonparametric
bootstrap. Assume that, based on a study, you obtained 6 individuals
who eat 100 g, 12 individuals who eat 150 g, 6 individuals who eat
170 g and 6 individuals who eat 200 g of ground beef. You want to
use a nonparametric bootstrap to derive uncertainty \cite{CULLEN-FREY-1999},
and then select samples from the empirical distribution.

<<sh22, echo=true, eval=true>>=
(x <- mcstoc(rempiricalD, type="U", outm=c('min','mean','max'), nvariates=30, 
              values=c(100,150,170,200),prob=c(6,12,6,6)))
mcstoc(rempiricalD,type="VU", values=x)
@

Printing the statistics of the 30 variates of \texttt{x} is of no
interest. Instead, we use the {}``\texttt{outm}'' option, which
allows us to specify which output we want ({}``\texttt{none}'' for
none, {}``\texttt{each}'', the default, for a series of statistics
for each variate, or, as in the example, a vector of functions that
are applied over all the 30 variates).

\subsection{Multivariate Nodes as a {}``Third Dimension'' for Multiple Options
in a Model}

The recycling rules in \texttt{mc2d} regarding the \texttt{nvariate}
dimension are as follows: the recycling will be done from \texttt{nvariates=1}
to \texttt{nvariates=n} with $n>1$. This allows you to use multivariates
nodes as a third dimension, in case you want to test various alternatives.

Assume, as in section \ref{sub:The-mcprobtree-function}, that the
distribution representing uncertainty in \texttt{conc} was not certain,
and that the microbiologists suggest that $conc\sim N\left(10,2\right)$
is possible, but that $conc\sim U\left(8,12\right)$ is also possible.
We can \emph{i}) build a {}``bivariate'' node reflecting these two
options; ii) transfer these options into the final risk estimate.
We obtain a bivariate node for the risk, one using the first hypothesis,
the second the second hypothesis.

<<sh23, echo=true>>=
conc1 <- mcstoc(rnorm, type="U", mean=10, sd=2)
conc2 <- mcstoc(runif, type="U", min=8, max=12)
conc <- mcdata(c(conc1,conc2),type="U",nvariates=2)

cook <- mcstoc(rempiricalD, type="V", values=c(1,1/5,1/50), prob=c(0.027,0.373,0.600))
serving <- mcstoc(rgamma,type="V",shape=3.93,rate=0.0806)
expo <- conc * cook * serving
dose <- mcstoc(rpois,type="VU",nvariates=2,lambda=expo)
r <- mcstoc(runif,type="U",min=0.0005,max=0.0015)
risk <- 1-(1-r)^dose
EC5 <- mc(conc,cook,serving,expo,dose,r,risk)
summary(EC5)
@

(Do not forget to transfer the number of variates you want in \texttt{mcstoc}...
(see the definition of \texttt{dose}). \texttt{mc2d} cannot guess...)

\subsection{Multivariate Nodes as a {}``Third Dimension'' for Multiple Vectors/Contaminants}

The recycling rules in \texttt{mc2d} also allow you to use multivariate
nodes as a third dimension for multiple vectors/Contaminants.

Assume in our ground beef example that we have two contaminants: one
has a mean concentration that follows an uncertainty distribution
$conc\sim N\left(10,2\right)$, the second one follows $conc\sim N\left(14,2\right)$.
We can \emph{i}) build a {}``bivariate'' node reflecting these two
concentrations%
\footnote{Note that we could simulate a correlation between both contaminants
using a multivariate normal distribution.%
} ; \emph{ii}) transfer these options into the final dose; \emph{iii})
sum the dose over the variates (using \texttt{mcapply}). The behavior
of contaminants is transferred in the model.

<<sh24, echo=true>>=
mconc <- mcdata(c(10,14), type="0", nvariates=2)
conc <- mcstoc(rnorm, nvariates=2, type="U", mean=mconc, sd=2)

cook <- mcstoc(rempiricalD, type="V", values=c(1,1/5,1/50), prob=c(0.027,0.373,0.600))
serving <- mcstoc(rgamma,type="V",shape=3.93,rate=0.0806)
expo <- conc * cook * serving
dose <- mcstoc(rpois,type="VU",nvariates=2,lambda=expo)
dosetot <- mcapply(dose, margin="variates", fun=sum)
r <- mcstoc(runif,type="U",min=0.0005,max=0.0015)
risk <- 1-(1-r)^dosetot
EC6 <- mc(conc,cook,serving,expo,dose,dosetot,r,risk)
summary(EC6)
@

As a conclusion, this \textquotedbl{}third\textquotedbl{} dimension
is highly flexible...

\section{Another Example: A QRA of Waterborne Cryptosporidiosis in France}

This example is adapted from \cite{POUILLOT-2004}. The aim is to
evaluate the risk of infection with \emph{Cryptosporidium} \emph{parvum}
from consumption of tap water, given that $n$ oocysts /100 l. have
been observed in a storage reservoir.

\subsection{Tap Water Consumption Model}

<<Crypto1, echo=false, eval=true>>=
library(mc2d)
inca <- structure(c(0, 22.08, 60, 64.4, 72, 82.8, 90, 96, 100, 110, 120,  137.5, 144, 150, 160, 162.5, 165, 180, 182.5, 184, 192, 192.5,  200, 216, 220, 225, 230, 240, 250, 264, 270, 276, 288, 290, 300,  304, 312.8, 320, 322, 325, 330, 336, 340, 350, 360, 370, 375,  380, 384, 390, 400, 414, 420, 425, 430, 432, 432.5, 440, 442,  450, 456, 460, 460.8, 464, 470, 470.4, 480, 490, 500, 504, 510,  510.4, 516, 520, 525, 525.6, 528, 530, 540, 544, 550, 552, 560,  562, 565, 570, 576, 580, 582.5, 584, 585.6, 590, 596, 600, 606,  610, 614, 620, 625, 630, 635.4, 640, 648, 650, 656.2, 660, 664.4,  670, 670.4, 672, 675, 680, 682, 690, 696, 700, 710, 716, 720,  730, 730.4, 740, 744, 750, 756, 760, 774.8, 780, 784, 792, 796,  800, 810, 820, 828, 830, 840, 850, 850.4, 860, 864, 866.4, 870,  880, 890, 900, 908, 910, 915.2, 920, 930, 936, 950, 960, 970,  980, 984, 986.4, 990, 996, 1000, 1015.2, 1020, 1028, 1032, 1036,  1040, 1042, 1050, 1070, 1075, 1078.8, 1080, 1090, 1096, 1100,  1110, 1120, 1126.4, 1128, 1130, 1140, 1148, 1150, 1152, 1160,  1170, 1175, 1176.2, 1190, 1196, 1200, 1214, 1220, 1230, 1240,  1248, 1250, 1260, 1276, 1280, 1290, 1296, 1300, 1320, 1322, 1330,  1340, 1350, 1360, 1370, 1386.4, 1400, 1410, 1414, 1420, 1430,  1440, 1446, 1450, 1460, 1480, 1500, 1520, 1530, 1550, 1560, 1600,  1650, 1680, 1696, 1700, 1710, 1720, 1750, 1760, 1800, 1830, 1840,  1850, 1900, 1920, 1936, 1954, 1980, 1990, 2000, 2014, 2050, 2100,  2200, 2220, 2248, 2250, 2276, 2300, 2310, 2340, 2400, 2550, 2568,  2700, 2720, 2784, 2820, 2876, 3000, 3100, 3108, 3200, 2578, 7,  1, 8, 14, 3, 1, 1, 10, 1, 250, 1, 2, 120, 8, 6, 1, 5, 3, 12,  5, 5, 375, 2, 8, 7, 41, 408, 53, 4, 24, 7, 3, 2, 217, 1, 1, 44,  9, 1, 31, 1, 1, 17, 294, 5, 3, 9, 3, 12, 525, 5, 23, 1, 3, 4,  1, 28, 3, 154, 2, 5, 1, 2, 6, 1, 299, 8, 148, 1, 2, 1, 1, 8,  3, 1, 2, 14, 20, 1, 18, 2, 20, 6, 1, 8, 2, 8, 1, 1, 1, 4, 1,  487, 3, 5, 1, 7, 1, 5, 1, 24, 3, 17, 1, 42, 1, 2, 1, 1, 1, 16,  1, 3, 1, 30, 4, 1, 183, 4, 1, 5, 1, 141, 1, 14, 1, 12, 1, 2,  1, 206, 6, 2, 1, 4, 92, 10, 1, 5, 1, 3, 5, 5, 2, 87, 1, 1, 1,  5, 5, 4, 4, 78, 1, 3, 2, 1, 16, 1, 133, 1, 5, 1, 1, 1, 4, 1,  43, 1, 1, 1, 30, 1, 1, 7, 2, 6, 1, 1, 3, 3, 1, 10, 1, 5, 1, 1,  1, 1, 1, 159, 2, 1, 1, 10, 1, 16, 4, 2, 5, 3, 1, 3, 11, 4, 1,  2, 12, 5, 1, 1, 44, 3, 2, 1, 2, 17, 1, 4, 1, 1, 17, 1, 1, 3,  4, 18, 14, 4, 1, 2, 1, 1, 1, 2, 12, 1, 2, 1, 1, 1, 1, 1, 3, 1,  20, 1, 1, 1, 7, 1, 1, 3, 1, 2, 2, 1, 6, 1, 1, 1, 1, 1, 1, 1,  2, 1, 1, 2), .Dim = c(270L, 2L))
inca <- rep(inca[,1],inca[,2])/1000 
@

We have raw data of daily consumption of tap water from 1,180 tap
water consumers (var \texttt{inca}, see Figure \ref{fig:Function-inca}).
We could choose to use this empirical distribution to evaluate the
variability in the tap water consumption:

%
\begin{figure}
\caption{\label{fig:Function-inca}Histogram of daily tap water intake}

<<Crypto2, echo=false, eval=true, fig=true>>=
hist(inca, xlab="l/day",freq=FALSE,main="") 
@
\end{figure}

<<Crypto3, echo=true>>=
ndvar(1001);ndunc(1001)
mcstoc(rempiricalD,type="V",values=inca)
@

but we will use the \textquotedbl{}\texttt{fitdistrplus}\textquotedbl{}
library. \texttt{inca} includes a lot of \texttt{0} nodes, corresponding
to days when individuals do not drink tap water (possibly they drink
bottled water on those days). We could try a mixture of distributions,
with \textquotedbl{}\texttt{0}\textquotedbl{}and \textquotedbl{}non-\texttt{0}\textquotedbl{}
data.

<<Crypto4, echo=true, figure=false>>=
library(fitdistrplus) 
pzero <- sum(inca==0)/length(inca)
inca_non_0 <- inca[inca!=0] 
descdist(inca_non_0)
@

%
\begin{figure}
\caption{\label{fig:Function-descdist}Graph from the \texttt{descdist} function. }

<<Crypto5, echo=false, eval=true, fig=true>>=
descdist(inca_non_0)
@
\end{figure}

Following the \texttt{descdist} function (See figure \ref{fig:Function-descdist}),
let us try the lognormal distribution.

<<Crypto6, echo=true, figure=false>>=
Adj_water <- fitdist(inca_non_0,"lnorm",method="mle")
meanlog <- Adj_water$est[1]
sdlog   <- Adj_water$est[2]
summary(Adj_water)
plot(Adj_water)
@

%
\begin{figure}
\caption{\label{fig:Function-mledist}Graph from the \texttt{fitdist} function. }

<<Crypto7, echo=false, eval=true, fig=true>>=
plot(Adj_water)
@
\end{figure}

Not so bad (See Figure \ref{fig:Function-mledist}), and better than
a gamma distribution (results not shown). We can now rebuild our mixture.
We could consider uncertainty around the maximum likelihood estimates
using the \texttt{bootdist} function of the \texttt{fitdistrplus}
package, using something like:

<<Crypto7bis, echo=true, eval=false>>=
Boot <- bootdist(Ajust_lnorm, bootmethod="param", niter=ndunc())
Mean_conso <- mcdata(Boot$estim$meanlog, type="U")
Sd_conso <- mcdata(Boot$estim$sdlog, type="U")
conso1 <- mcstoc(rlnorm, type="VU", meanlog= Mean_conso, sdlog= Sd_conso)
@

But for simplicity, we will not consider uncertainty around the estimates.

We will use the \texttt{mcprobtree} function to construct a mixture
of \textquotedbl{}0\textquotedbl{}and \textquotedbl{}non-0\textquotedbl{}
distributions:

<<Crypto8, echo=true, figure=false>>=
conso0 <- mcdata(0,type="V")
conso1 <- mcstoc(rlnorm, type="V", meanlog=meanlog, sdlog=sdlog)
v <- mcprobtree(c(pzero,1-pzero), list("0"=conso0,"1"=conso1), type = "V")
summary(v)
@

\subsection{The Dose-Response Model}

We propose a boostrap from data (\texttt{datDR}) derived from \cite{CHAPPELL-1996}.
We first define a function \textquotedbl{}\texttt{DR}\textquotedbl{}
with an \texttt{n} argument for the size of the sample to draw. This
function may then be used in a \texttt{mcstoc} function:

<<Crypto9, echo=true, figure=false>>=
datDR <- list(   dose=c(30,100,300,500,1000,10000,100000,1000000),
                 pi=c(2,4,2,5,2,3,1,1),
                 ni=c(5,8,3,6,2,3,1,1))

estDR <- function(pos,ref){
   -glm(cbind(ref$ni-pos,pos) ~ ref$dose + 0, binomial(link="log"))$coefficients}

ml <- 1-exp(-estDR(datDR$pi, datDR) * datDR$dose)

DR <- function(n){
   boot <- matrix(rbinom(length(datDR$dose)*n,datDR$ni,ml),nrow=length(datDR$dose))
   apply(boot,2,estDR,ref=datDR)}
r <- mcstoc(DR, type="U")
summary(r)
@

\subsection{The Model}

Deriving the final model is straightforward. We construct the \texttt{mcnode}
corresponding to the recovery rate (Uncertainty,\texttt{ Rr}), the
probability for an oocyst to be infective (Variability, \texttt{w}):

<<Crypto10, echo=true, figure=false>>=
Rr <- mcstoc(rbeta, type="U", shape1=2.65, shape2=3.64)
w <- mcstoc(rbeta, type="V", shape1=2.6, shape2=3.4) 
@

Given that $O_{o}=2$ oocysts are observed in 100 l of water, the
expected number of oocysts in the sample is \texttt{l}:

<<Crypto11, echo=true, figure=false>>=
Oo <- 2
l <- (Oo + mcstoc(rnbinom, type="U", size=Oo+1, prob=Rr))/100
@

The expected number of oocysts drunk by the individuals is \texttt{Or}
and the risk ($\times10000$) is estimated by:

<<Crypto12, echo=true, figure=false>>=
Or <- l * v * w
P <- 10000 * (1-exp(-r*Or))
summary(P)
@

This result can be compared (roughly since there is some differences
in the modelsvariability) to the results shown in Table 2 in \cite{POUILLOT-2004}.

Improvement: the results for $O_{o}=\left\{ 0,1,2,5,10,20,50,100,1000\right\} $
can be obtained in one step using:

<<Crypto13, echo=true, eval=false, figure=false>>=
Oo <- mcdata(c(0,1,2,5,10,20,50,100,1000),type="0",nvariates=9) 
@

\section*{As a Conclusion}

We think and hope that {}``\texttt{mc2d}'' could help risk assessors
to constuct and analyse their models, and that it may help in developing
\textquotedbl{}two-dimensional\textquotedbl{} simulations. Nevertheless,
\textquotedbl{}\texttt{mc2d}\textquotedbl{} is currently under development:

\noindent \begin{center}
\emph{CHECK YOUR MODEL CAREFULLY AND EXAMINE RESULTS TO DETECT BUGS}
\par\end{center}

and, if you would like to improve it, join us at

\noindent \begin{center}
\emph{\url{http://riskassessment.r-forge.r-project.org/}}
\par\end{center}

Please refer any comments or bugs to \url{rpouillot@yahoo.fr}.

\bibliographystyle{plain}
\bibliography{Bibtex}

\end{document}