### 2010-05-28 pkg/mc2d/inst/doc/mc2dLmEnglish.rnw
Version with 2 vignettes
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/riskassessment@101 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% mc2d: Tools for Two-Dimensional Monte-Carlo Simulations
%%%
%%%         New TP V.2
%%%

\documentclass[10pt, english]{article}

%\usepackage[letterpaper]{geometry}
\usepackage[T1]{fontenc}
\usepackage[latin9]{inputenc}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage[scale=0.8, centering]{geometry}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% User specified LaTeX commands.
% sweave commands for vignette
%\VignetteIndexEntry{Case study: L. monocytogenes in cold-smoked salmon}
%\VignettePackage{mc2d}
%\VignetteDepends{mvtnorm,fitdistrplus}

%\makeatother

\begin{document}

% Command for Sweave: figure = .5 the text width
\setkeys{Gin}{width=0.5\textwidth}

\title{Case study: \emph{L.~monocytogenes} in cold-smoked salmon}
\author{R. POUILLOT, M.-L. DELIGNETTE-MULLER, M. CORNU}

\maketitle

The objective of this case study is to assess the risk of invasive listeriosis from consumption
of cold-smoked salmon in France. 
The process of interest lays from the end of the production line in the factory, when the cold-smoked 
salmon is vacuum-packed, to the consumption.

The data and the model are adapted to illustrate the use 
of \texttt{mc2d}: the results will not and \emph{should not} be interpreted as an assessment
of the actual risk of listeriosis from consumption of cold-smoked salmon. 
Interested readers could refer to \cite{POUILLOT-2007} and \cite{POUILLOT-2009} for a complete
risk assessment on that issue.

The model will be developed in a first section, without considering variability or uncertainty (deterministic model).
Variability will then be introduced in a second section, and a last section will consider variability 
and a part of the data uncertainty.

\section{The Model}

In this section, no variability nor uncertainty is considered.
We assess the final level of \emph{L.~monocytogenes} in the product,
the exposure and the risk of invasive listeriosis for an ``average'' individual
of the ``healthy'' French population\footnote{Yes, it makes no sense, but it will help us introducing
smoothly the model.}.

During the logistic, the retail and the home step,
a bacterial growth is modeled considering \emph{i}) the fluctuating temperature
during the various stages and; \emph{ii}) the bacterial competition with the food flora.
We use the models developed and/or used in \cite{POUILLOT-2007}. The data are adapted from
\cite{POUILLOT-2007} and \cite{DELIGNETTE-2006}:

\begin{itemize}
  \item The DMS model predicts the bacterial growth during a stage of duration $d$,
  when the temperature is fluctuating, with an intra-stage average temperature $m_{T}$
  and an intra-stage standard deviation of the temperature $s_{T}$. It is written:
  %
  \begin{equation}
  N_{1}=
  \min\left(
    N_{0} +
    \frac{\mu_{ref}}{\ln(10)} \times d \times 
    \frac{\left(s_{T}^{2}+\left(m_{T}-T_{min}\right)^{2}\right)}{\left(T_{ref}-T_{min}\right)^{2}},
    \: N_{max}\right) \label{DMS}
  \end{equation}
  %
  if $m_{T}>T_{min}$, with $N_{1}$ the $\log_{10}$ concentration of bacteria
  ($\log_{10}$ (CFU/g)) in the product at the end of the stage,
  $N_{0}$ the $\log_{10}$ concentration of bacteria ($\log_{10}$ (CFU/g))
  in the product at the beginning of the stage, $\mu_{ref}$ the specific growth
  rate ($\text{day}^{-1}$) at a reference temperature $T_{ref}$ (\textdegree{}C),
   $T_{min}$ the minimal temperature (\textdegree{}C) of growth
  and $N_{max}$   the maximum achievable concentration in the product ($\log_{10}$ (CFU/g)).
  If $m_{T} \leq T_{min}$, $N_{1}=N_{0}$.
  \item We will use $T_{ref}=25$\textdegree{}C.
  We have in this section $N_{max}=7.27 \log_{10}(\text{CFU/g})$;
  \item The model for \emph{L.~monocytogenes} uses $\mu_{ref,Lm}=6.2\: \text{day}^{-1}$ and
  $T_{min,Lm}=-2.9$\textdegree{}C;
  \item  The same model is used for the food flora, with $\mu_{ref,ff}=4.1\: \text{day}^{-1}$ and
  $T_{min,ff}=-4.5$\textdegree{}C;
  \item The growth model for the bacterial competition consider the Jameson effect,
  i.e.~consider that the bacterial growth of \emph{L.~monocytogenes and} the growth of the food flora
  are stopped as soon as one population reaches $N_{max}$.
\end{itemize}
  In practice, one will evaluate $d_{Lm}$ and $d_{ff}$, the time needed for \emph{L.~monocytogenes}
  or the food flora to reach $N_{max}$, respectively, and model a growth for the
  given stage during an effective duration of $\min(d,d_{Lm},d_{ff})$.
  The time needed to reach $N_{max}$ is evaluated by inverting \eqref{DMS}:
  %
  \[
  d_{\left(N_{1}=N_{max}\right)}=
  \left(N_{max}-N_{0}\right)\times
    \frac{\ln(10)}{\mu_{ref}} \times
    \frac{\left(T_{ref}-T_{min}\right)^{2}}{\left(s_{T}^{2}+\left(m_{T}-T_{min}\right)^{2}\right)}
  \]
  %

The other assumptions are:
\begin{itemize}
  \item  A cold-smoked salmon package is homogeneously contaminated with \emph{L.~monocytogenes}
  at the end of the production at a level of $0.1$ CFU/g;
  \item  The food flora level at the end of the production is $10^{2.78}$ CFU/g;
  \item  The time-temperature profile is:

    \begin{itemize}
    \item  1.1 days at an average temperature of 3.2\textdegree{}C from the factory to the retail (logistic step),
    with an intra-stage standard deviation of the temperature of 2.1 \textdegree{}C;
    \item  4.7 days at an average temperature of 5.5\textdegree{}C at retail
    with an intra-stage standard deviation of the temperature of 1.0 \textdegree{}C;
    \item  4.3 days at an average temperature of 8.2\textdegree{}C in the consumer's home
    with an intra-stage standard deviation of the temperature of 2.0 \textdegree{}C;
    \end{itemize}
  \item  An healthy, non elderly, non pregnant individual eats 35g of this product;
  \item  The individual dose-response model for this population is a one hit model
  \[
  \Pr(\text{Illness} \mid D)=1-(1-r)^{D}
  \]
  with $r=4.7\times10^{-14}$ for an individual from this healthy sub-population.
  The populational dose-response that evaluates the mean risk for
  a population exposed to food where the number of bacteria follows a Poisson distribution of
  mean paramer $D$ is the exponential dose-response
  \[
  \Pr(\text{Illness} \mid D)=1-\exp(r\times D)
  \]
\end{itemize}

The question is ``What is the risk for this `average' individual?''. 
One way to write this model is as following:

%Specify the width opf the output
<<sh0, echo=false, eval=true>>=
options("width"=100,"digits"=3)
set.seed(666)
@

<<sh1, echo=true, eval=true>>=
Nmax <- 7.3
murefLm <- 6.2; TminLm <- -2.9

murefFF <- 4.1; TminFF <- -4.5

Lm0 <- log10(1); FF0 <- 2.78

d1 <- 1.1; mT1 <- 3.2; sdT1 <- 2.1

d2 <- 4.7; mT2 <- 5.5; sdT2 <- 1.0

d3 <- 4.3; mT3 <- 8.2; sdT3 <- 2.0

conso <- 35
r <- 4.7e-14

modGrowth <- function(duration, mTemp, sdTemp,
                      N0Lm, murefLm, TminLm,
                      N0FF, murefFF, TminFF,
                      Nmax, Tref=25) {
  N0Lm <- pmin(N0Lm, Nmax)
  N0FF <- pmin(N0FF, Nmax)
  dLm <- (Nmax-N0Lm) * log(10)/murefLm * (Tref-TminLm)^2 / (sdTemp^2 + (mTemp-TminLm)^2)
  dLm <- ifelse(mTemp < TminLm & N0Lm!=Nmax, Inf, dLm)
  dFF <- (Nmax-N0FF) * log(10)/murefFF * (Tref-TminFF)^2 / (sdTemp^2 + (mTemp-TminFF)^2)
  dFF <- ifelse(mTemp < TminFF & N0FF!=Nmax, Inf, dFF)
  realDuration <- pmin(duration, dLm , dFF)
  xLm <- N0Lm + (mTemp > TminLm) * murefLm/log(10) * (sdTemp^2 + (mTemp - TminLm)^2) /  ((Tref - TminLm)^2) * realDuration
  xFF <- N0FF + (mTemp > TminFF) * murefFF/log(10) * (sdTemp^2 + (mTemp - TminFF)^2) /  ((Tref - TminFF)^2) * realDuration
  return(list(xLm = xLm, xFF=xFF))}

x1 <- modGrowth(d1, mT1, sdT1,
                Lm0, murefLm, TminLm,
                FF0, murefFF, TminFF,
                Nmax)
x2 <- modGrowth(d2, mT2, sdT2,
                x1$xLm, murefLm, TminLm,
                x1$xFF, murefFF, TminFF,
                Nmax)
x3 <- modGrowth(d3, mT3, sdT3,
                x2$xLm, murefLm, TminLm,
                x2$xFF, murefFF, TminFF,
                Nmax)
x3
conta <-10^x3$xLm
conta
expo <- conso * conta
expo
risk <- 1 - (1 - r)^expo
risk
@

\texttt{modGrowth} is a convenient function for the growth model. Within this function
\texttt{dLm} is the time needed for \emph{L.~monocytogenes} to reach \texttt{Nmax},
\texttt{dFF} is the time needed for the food flora to reach \texttt{Nmax} and,
\texttt{realDuration} is the effective time of growth during the stage. Note that: 
\begin{itemize}
\item this function is ``vectorized'', meaning that it can deal with a vector for any of its parameters, returning consequently 
a vector. This is a strength of R, notably for Monte-Carlo simulations, but it requests a bit of knowledge on
the way to code the functions. As an example: \texttt{pmin}, a function that takes one or more vectors 
as arguments and return a single vector giving the ``parallel'' minima of the vectors  is used instead
of the more classical function \texttt{min} function, that would return the maximum or minimum of \emph{all} the values.
Another example is the use of the \texttt{ifelse} instead of \texttt{if};
\item it is also written to handle \emph{all} specific cases that could occur in the Monte-Carlo simulation, such as 
$N_{0} \geq N_{max}$ or $m_{T} \leq T_{min}$ or both, for any or both bacterial populations.    
\end{itemize}
\texttt{x1}, \texttt{x2} and \texttt{x3} are the bacterial concentrations at the end of the logistic, the retail and the home step, respectively.

\section{Including Variability}

We now specify now some variability distributions for some inputs, following \cite{DELIGNETTE-2006} and
\cite{POUILLOT-2007}.
We first have to call the needed libraries, and define the desired number of iterations:

<<shCallLibrary, echo=true, eval=true>>=
library(fitdistrplus)
library(mc2d)
ndvar(10001)
@

\subsection{Specifying Variability Distribution}
\subsubsection{Initial Contamination}
For the initial contamination levels in \emph{L.~monocytogenes}, 
we have a set of 62 enumeration data from a representative sample of
packages of cold smoked salmon positive in detection: 43 samples have less than $0.2$ CFU/g,
7 samples have $0.2$ CFU/g, 4 samples have $0.4$ CFU/g, 2 samples have $0.6$ CFU/g,
and the other values are $0.3$, $1.0$, $1.6$, $2.4$, $5.4$ and $7.0$ CFU/g \cite{POUILLOT-2007}. 
We will use
the \texttt{fitdistrplus} package to fit a normal distribution on the $\log_{10}$ of
these values, taking into account the censored values.
Using the fitted parameters, we model thereafter these initial concentrations
in contaminated packages through a normal distribution truncated\footnote{so that at least one CFU is included in one 100g package}
on $[-2,\infty)\: \log_{10}$ (CFU/g).

For the food flora, we use the distribution proposed by \cite{DELIGNETTE-2006}, 
$N_{0ff} \sim N(2.78,1.14)$.

<<shLm0FF0V, echo=true, eval=true>>=
dataC <- data.frame(
  left =c(rep(NA,43), rep(.2,7),.3,rep(.4,4),1,1.6,.6,.6,2.4,5.4,7),
  right=c(rep(0.2,43),rep(.2,7),.3,rep(.4,4),1,1.6,.6,.6,2.4,5.4,7)
  )
fit <- fitdistcens(log10(dataC), "norm")
fit
Lm0V <- mcstoc(rnorm, mean = fit$est["mean"], sd = fit$est["sd"], rtrunc=TRUE, linf=-2)
FF0V <- mcstoc(rnorm, mean=2.78, sd=1.14)
@
Note that, by default, the type of alea that is modeled is ``variability'' (\texttt{type="V"}).

\subsubsection{Growth Parameters}
Distributions are derived from \cite{DELIGNETTE-2006}: 
\begin{itemize}
\item $N_{max}$ follows a normal distribution with mean 7.27 $\log_{10}$ CFU/g
and standard deviation 0.86 $\log_{10}$ CFU/g;
\item The specific growth rate at the reference temperature of 25\textdegree{}C
for \emph{L.~monocytogenes} follows a normal distribution with mean 6.24 $\text{day}^{-1}$ and standard
deviation 0.75 $\text{day}^{-1}$ truncated on $[0,\infty)$. The minimal growth temperature follows a normal distribution with
mean -2.86\textdegree{}C and standard deviation 1.93\textdegree{}C;
\item The specific growth rate at the reference temperature of 25\textdegree{}C
for the food flora follows a normal distribution with mean 4.12 $\text{day}^{-1}$ and standard
deviation 1.97 $\text{day}^{-1}$ truncated on $[0,\infty)$. The minimal growth temperature follows a normal distribution with
mean -4.52\textdegree{}C and standard deviation 7.6\textdegree{}C.
\end{itemize}

<<shGrowtV, echo=true, eval=true>>=

NmaxV    <- mcstoc(rnorm, mean=7.27, sd = 0.86)

murefLmV <- mcstoc(rnorm, mean = 6.24, sd = 0.75, rtrunc=TRUE, linf=0)
TminLmV  <- mcstoc(rnorm, mean = -2.86, sd = 1.93)

murefFFV <- mcstoc(rnorm, mean = 4.12, sd = 1.97, rtrunc=TRUE, linf=0)
TminFFV <-  mcstoc(rnorm, mean = -4.52, sd = 7.66)
@

\subsubsection{Time-Temperature Profiles}
The time temperature profiles in the three steps are modelled using
the distribution provided in the table \ref{tab:4} (adapted from \cite{POUILLOT-2007} from representative
data from France)\footnote{$\Gamma$ is the Gamma distribution parameterized as $\Gamma(shape,\: scale)$.
The Exponential($x$) ditribution is the exponential distribution with mean $x$.}. We
assume a shelf life of 28 days. A simple way to model this shelf life will
be to have $d_{1}+d_{2}+d_{3}\leq28$ days, with $d_{1}$ the duration of the logistic stage,
$d_{2}$ the duration of the retail stage and $d_{3}$ the duration of the consumer stage\footnote{See
the code for a way to model this shelf life using truncated distributions.};

\begin{table}
 \caption{ \label{tab:4}Time Temperature Profiles}

 \centering{}\begin{tabular}{|c|c|c|c|}
\hline 
 Stage &  Mean Temperature (\textdegree{}C) & Intra-Stage Variance of T (\textdegree{}C) &  time (days)%
\tabularnewline
\hline
\hline 
 logistic &  normal(3.2, 2.2) truncated on {[}-3;25{]} & $\Gamma(1.16, 4.61)$ &  Exponential(1.1)\tabularnewline
\hline 
 retail &  normal(5.5, 2.2) truncated on {[}-3;25{]} & $\Gamma(0.65, 2.09)$ &  Exponential(4.7)\tabularnewline
\hline 
 consumer &  normal(8.2, 3.8) truncated on {[}-3; 25{]} & $\Gamma(0.35, 19.7)$  &  Exponential(4.3)\tabularnewline
\hline
\end{tabular}
\end{table}
%

<<shTtV, echo=true, eval=true>>=
d1V <- mcstoc(rexp, rate = 1/1.1)
mT1V <- mcstoc(rnorm, mean = 3.2, sd = 2.2, rtrunc = TRUE, linf = -3, lsup = 25)
sdT1V <- sqrt(mcstoc(rgamma, shape = 1.16, scale=4.61))

d2V <- mcstoc(rexp, rate = 1/4.7, rtrunc=TRUE, lsup=28-d1)
mT2V <- mcstoc(rnorm, mean = 5.5, sd = 2.2, rtrunc = TRUE, linf = -3, lsup = 25)
sdT2V <- sqrt(mcstoc(rgamma, shape = 0.65, scale=2.09))

d3V <- mcstoc(rexp, rate = 1/4.3, rtrunc=TRUE, lsup=28-(d1+d2))
mT3V <- mcstoc(rnorm, mean = 8.2, sd = 3.8, rtrunc = TRUE, linf = -3, lsup = 25)
sdT3V <- sqrt(mcstoc(rgamma, shape = 0.35, scale=19.7))
@

\subsubsection{Serving Size}
As for the serving size, we consider, from observed data, a discrete empirical distribution with values \cite{POUILLOT-2007}:
$V$=\{10, 12, 19, 20, 30, 34, 40, 50, 60, 67.5, 80, 100, 250\} grams, observed
$F$=\{11, 1, 1 ,29, 12, 1, 41, 4, 4, 1, 4, 1, 1\} time, respectively.
%
<<shConsoV, echo=true, eval=true>>=
consoV <- mcstoc(rempiricalD, values = c(10, 12, 19, 20, 30, 34, 40, 50, 60, 67.5, 80, 100, 250), prob = c(11, 1, 1, 29, 12, 1, 41, 4, 4, 1, 4, 1, 1))
@

\subsection{Applying the Model}
The model may then be evaluated straightforwardly:

<<shModelV, echo=true, eval=true>>=
r <- mcdata(4.7e-14, type = "0")
x1V <- modGrowth(d1V, mT1V, sdT1V,
                Lm0V, murefLmV, TminLmV,
                FF0V, murefFFV, TminFFV,
                NmaxV)
x2V <- modGrowth(d2V, mT2V, sdT2V,
                x1V$xLm, murefLmV, TminLmV,
                x1V$xFF, murefFFV, TminFFV,
                NmaxV)
x3V <- modGrowth(d3V, mT3V, sdT3V,
                x2V$xLm, murefLmV, TminLmV,
                x2V$xFF, murefFFV, TminFFV,
                NmaxV)

contaV <-10^x3V$xLm
expoV <- consoV * contaV
riskV <- 1 - exp(-r * expoV )
Lm1 <- mc(Lm0V, FF0V, NmaxV, murefLmV, TminLmV, murefFFV, TminFFV,
          d1V, mT1V, sdT1V, d2V, mT2V, sdT2V, d3V, mT3V, sdT3V,
          consoV, r, contaV, expoV, riskV)
Lm1
sLm1 <- mc(contaV=Lm1$contaV, expoV=Lm1$expoV, riskV=Lm1$riskV)
summary(sLm1, probs = c(0, 0.5, 0.75, 0.95, 1))
@
\texttt{Lm1} is a \texttt{mc} object that contains all the parameters and outputs.
We extract some of these outputs in \texttt{sLm1} to provide a short summary.

\subsection{Final Estimate}
If 6.5\% of cold-smoked salmon package are contaminated,
if 49,090,000 Frenchs are part of the ``non susceptible''
population and if, on average, those people consume some smoked salmon
$6.4$ times per year, the expected number of cases of listeriosis
from consumption of cold smoked salmon in this population is estimated through:

<<sh4, echo=true, eval=true>>=
meanRisk <-  mcapply(riskV,"var",mean)
expectedN <- round(0.065 * unmc(meanRisk) * 6.4 * 49090000)
expectedN
@

\section{Including (a Part of the) Uncertainty}
We eventually include both variability and uncertainty in the model.
For this example, we will only consider the uncertainty linked to the initial contamination,
the growth parameters and the prevalence.

\subsection{Specifying Uncertainty}
\subsubsection{Initial Contamination}

The uncertainty surrounding the initial contamination levels of the \emph{L.~monocytogenes}
will be modeled using a bootstrap procedure, obtained straightforwardly with the help of the \texttt{fitdistrplus}
package and its \texttt{bootdistcens} function. Before this, we define the number of iterations needed in 
the uncertainty dimension.
%
<<sh8, echo=true, eval=true>>=
ndunc(101)
bootLm0 <- bootdistcens(fit, niter=ndunc())
MLm0 <- mcdata(bootLm0$est$mean,type="U")
SLm0 <- mcdata(bootLm0$est$sd,type="U")
Lm0VU <- mcstoc(rnorm, type="VU", mean=MLm0, sd=SLm0, rtrunc=TRUE, linf=-2)
@

In order to consider uncertainty for the food flora initial contamination, we have, from \cite{DELIGNETTE-2006},
a set of uncertain hyperparameters, $M_{N0ff}$ and $\sigma_{N0ff}$, that are used
as parameters for the uncertain \emph{and} variable parameter $N_{0ff}$:
\begin{eqnarray*}
  N_{0ff} & \sim & N(M_{N0ff},\sigma_{N0ff})\\
  M_{N0ff} & \sim & N(2.78,0.265)\\
  \ln(\sigma_{N0ff}) & \sim & N(0.114,0.172)
\end{eqnarray*}

This hierarchical simulation is written with \texttt{mc2d}:

<<shFF0VU, echo=true, eval=true>>=
MLm0FF <- mcstoc(rnorm, type="U", mean=2.78, sd=0.265)
SLm0FF <- mcstoc(rlnorm, type="U", meanlog=0.114, sdlog=0.172)
FF0VU <- mcstoc(rnorm, type="VU", mean=MLm0FF, sd=SLm0FF)
@
\subsubsection{Growth Parameters}
The uncertainty around $\mu_{ref,Lm}$, $T_{min,Lm}$, $\mu_{ref,ff}$, $T_{min,ff}$
and $N_{max}$ are modeled similarly through the
specification of hyperparameters \cite{DELIGNETTE-2006}\footnote{Note
that there was a typo in \cite{DELIGNETTE-2006} that lead to an error in
\cite{POUILLOT-2007}: the standard-error for $\ln(\sigma_{\mu ref,Lm})$ is $1.03$
and not $-1.03$ as written in \cite{DELIGNETTE-2006}. We will use here the correct value.}:
\begin{eqnarray*}
  \mu_{ref,Lm} & \sim & N(M_{\mu ref,Lm},\sigma_{\mu ref,Lm})\\
  M_{\mu ref,Lm} & \sim & \Gamma(shape: 69.7, scale: 0.0896)\\
  \ln(\sigma_{\mu ref,Lm}) & \sim & N(1.03,0.191)\\
\\
  T_{min,Lm} & \sim & N(M_{Tmin,Lm},\sigma_{Tmin,Lm})\\
  M_{Tmin,Lm} & \sim & N(-2.86,0.459)\\
  \ln(\sigma_{Tmin,Lm}) & \sim & N(0.638,0.208) \\
\\
  \mu_{ref,ff} & \sim & N(M_{\mu ref,ff},\sigma_{\mu ref,ff})\\
  M_{\mu ref,ff} & \sim & \Gamma(shape: 32.5, scale: 0.127)\\
  \ln(\sigma_{\mu ref,ff}) & \sim & N(-0.656,0.221)\\
\\
  T_{min,ff} & \sim & N(M_{Tmin,ff},\sigma_{Tmin,ff})\\
  M_{Tmin,ff} & \sim & N(-4.52,1.23)\\
  \ln(\sigma_{Tmin,ff}) & \sim & N(2.00,0.257) \\
\\
  N_{max} & \sim & N(M_{Nmax},\sigma_{Nmax})\\
  M_{Nmax} & \sim & N(7.27,0.276)\\
  \ln(\sigma_{Nmax}) & \sim & N(-0.172,0.218)
\end{eqnarray*}

with $\mu_{ref}>0$ and $T_{min}<25$. We simply translated the preceeding distributions:

<<sh5, echo=true, eval=true>>=
MmurefLm <- mcstoc(rgamma, type="U", shape=69.7, scale=0.0896)
SmurefLm <- mcstoc(rlnorm, type="U", meanlog = 1.03, sdlog = 0.191)
murefLmVU <- mcstoc(rnorm, type="VU", mean=MmurefLm, sd=SmurefLm, rtrunc=TRUE, linf=0)

MTminLm <- mcstoc(rnorm, type="U", mean=-2.86, sd=0.459)
STminLm <- mcstoc(rlnorm, type="U", meanlog = 0.638, sdlog = 0.208)
TminLmVU <- mcstoc(rnorm, type="VU", mean = MTminLm, sd = STminLm, rtrunc=TRUE, lsup=25)

MmurefFF <- mcstoc(rgamma, type="U", shape=32.5, scale=.127)
SmurefFF <- mcstoc(rlnorm, type="U", meanlog = -.656, sdlog = 0.221)
murefFFVU <- mcstoc(rnorm, type="VU", mean=MmurefFF, sd=SmurefFF, rtrunc=TRUE, linf=0)

MTminFF <- mcstoc(rnorm, type="U", mean=-4.52, sd=1.23)
STminFF <- mcstoc(rlnorm, type="U", meanlog = 2.00, sdlog = 0.257)
TminFFVU <- mcstoc(rnorm, type="VU", mean = MTminFF, sd = STminFF, rtrunc=TRUE, lsup=25)

MNmax <- mcstoc(rnorm, type="U", mean=7.27, sd=0.276)
SNmax <- mcstoc(rlnorm, type="U", meanlog = -0.172, sdlog = 0.218)
NmaxVU <- mcstoc(rnorm, type="VU", mean = MNmax, sd = SNmax)
@
%
\subsubsection{Prevalence}
The prevalence level of contaminated cold-smoked salmon packages
(6.5\%) was estimated from 41 positive packages out of 626 tested \cite{POUILLOT-2007}.
We assume a sensitivity and a specificity of the method of 100\%.
We model the data uncertainty around the true prevalence of contaminated
package using a bayesian reasonning, with a Beta(1, 1) distribution
as a prior. The number of expected cases may be estimated using:
<<shPrevU, echo=true, eval=true>>=
prevU <- mcstoc(rbeta,type="U", shape1=41+1, shape2=626-41+1)
@
\subsection{Applying the Model}
Applying the model is just a copy-paste from the previous version (+ we change the name of the parameters).

<<sh9, echo=true, eval=true>>=
x1VU <- modGrowth(d1V, mT1V, sdT1V,
                Lm0VU, murefLmVU, TminLmVU,
                FF0VU, murefFFVU, TminFFVU,
                NmaxVU)
x2VU <- modGrowth(d2V, mT2V, sdT2V,
                x1VU$xLm, murefLmVU, TminLmVU,
                x1VU$xFF, murefFFVU, TminFFVU,
                NmaxVU)
x3VU <- modGrowth(d3V, mT3V, sdT3V,
                x2VU$xLm, murefLmVU, TminLmVU,
                x2VU$xFF, murefFFVU, TminFFVU,
                NmaxVU)

contaVU <-10^x3VU$xLm
expoVU <- consoV * contaVU
riskVU <- 1 - exp(-r * expoVU)
Lm2 <- mc(Lm0VU, FF0VU, NmaxVU, murefLmVU, TminLmVU, murefFFVU, TminFFVU,
          d1V, mT1V, sdT1V, d2V, mT2V, sdT2V, d3V, mT3V, sdT3V,
          consoV, r, contaVU, expoVU, riskVU)
Lm2
sLm2 <- mc(contaVU=Lm2$contaVU, expoVU=Lm2$expoVU, riskVU=Lm2$riskVU)
summary(sLm2, probs = c(0, 0.5, 0.75, 0.95, 1))
@
The summary provides the estimate of the mean, the standard deviation, the minimum, the median \ldots and a 95\% credible interval.
The estimate is the median of the 101 values obtained in the uncertainty dimension. The credible interval lays
between the $2.5^{\text{th}}$ and the $97.5^{\text{th}}$ percentiles obtained in the uncertainty dimension.

\subsection{Final Estimate}
The uncertainty around the number of expected cases is estimated using:
<<sh7, echo=true, eval=true>>=
meanRiskU <-  mcapply(riskVU,"var",mean)
expectedNU <- round(prevU * meanRiskU * 6.4 * 49090000)
summary(expectedNU)
@

This is an estimate of the uncertainty around the number of cases linked to the uncertainty 
around the initial contamination, the bacterial growth parameter and
the sampling uncertainty for positive packages.
A lot of other uncertainties exist but are not considered here, notably 
the uncertainty around the dose-response model
and parameters. See \cite{POUILLOT-2007,POUILLOT-2009} for a complete analysis.
The study of the model through a Tornado chart in the variability dimension
leads to the Figure \ref{fig:TorLm}. 
It suggests a big impact of the growth rate of \emph{L.~monocytogenes}, of the storage duration
during the consumer step, and of the initial level of \emph{L.~monocytogenes}. 
The Tornado chart in the uncertainty dimension
leads to the Figure \ref{fig:TorLmU} and suggests the impact of the uncertainty around
$N_{max}$ on the mean risk, and thus the expected number of cases.

<<sh3, echo=true, eval=true, figure=false>>=
torn <- tornado(Lm2)
torn
tornunc <- tornadounc.mc(Lm2, quant=NULL)
tornunc
@
\texttt{quant=NULL} request an analysis only on the mean risk, and not on some qunatiles.
<<sh3b, echo=true, eval=false, figure=false>>=
plot(torn)
plot(tornunc, stat="mean risk")
@

\begin{figure}
\caption{\label{fig:TorLm}Tornado chart for the \emph{L.~monocytogenes} example (Variability).}
\begin{center}
<<sh3c, echo=false, eval=true, fig=true>>=
plot(torn)
@
\end{center}
\end{figure}

\begin{figure}
\caption{\label{fig:TorLmU}Tornado chart for the \emph{L.~monocytogenes} example (Uncertainty).}
\begin{center}
<<sh3d, echo=false, eval=true, fig=true>>=
plot(tornunc, stat="mean risk")
@
\end{center}
\end{figure}

As a conclusion, this example illustrates how predictive growth models may 
be implemented within \texttt{mc2d}\ldots

\bibliographystyle{plain}
\bibliography{docmcEnglish}

\end{document}